A performance assessment of evolutionary algorithms in volunteer computing environments: the importance of entropy

### 2017-04-16 EvoPar2017/paper.Rnw
Adds new figures to avoid deletion

\begin{document}
%\SweaveOpts{concordance=TRUE}

<<setup, cache=FALSE,echo=FALSE>>=
library("ggplot2")
library("RColorBrewer")
opts_chunk$set( dev        = c("pdf", "png") )
experiment.data <- data.frame(Experiment=character(),
                              milliseconds=character(),                           
                              IPs=character(), 
                              actualIPs=character(),
                              PUTs=character(),
                              actualPUTs=character(),
                              PUTratio = character(),
                              PUTsIP =character())

@@

%
%
\begin{figure}[htbp]
\centering
<<zipf, cache=FALSE,echo=FALSE,fig.height=6>>=
ggplot(ips.data,aes(x=idx,y=PUTs,group=Experiment,color=Experiment))+geom_point()+scale_y_log10()+ scale_colour_discrete(labels=c('base','graph1','graph2','reboot','cachereboot'))+scale_x_log10()
@ %def graph for IPs and puts
%
\caption{Log-log plot of ranked and normalized number of contributions (PUTs) per IP.\label{fig:zipf}}
\end{figure}
%
Every run had slightly different conditions, although in some cases it
was just a server reboot and a new round of announcements. Each run
batch included at least 30 completions, that is, evolution until the
solution was found. The number of runs was controlled by polling an
URL that indicated the number of runs made in the present batch so
far. Every volunteer contributed a certain number of individuals
(represented by the PUT HTTP request in Figure \ref{fig:system}) to the server. The ranked number
of contributions per unique IP is shown in Figure \ref{fig:zipf} in a
log-log scale. Besides following the usual Zipf's law found in our
previous work and thus having a very similar appearance for all 5 sets, it
is interesting to see how some sets have a number of contributions per
IP that are slightly superior to the rest, especially so in {\tt
  graph2}. The last two sets, {\tt reboot} and {\tt cachereboot}, also
show a regime in the middle of the graph which shows that some users
are contributing more than in other similar sets. In general, these
differences in behavior are due more to the specific users that show
up in these experiments than to general laws, but it should also be
noted that those behaviors can be encouraged by design.

@@

\section{Finding the keys to a good performance}
%
%
\begin{figure}[ht!b]
\centering
<<minutedata, cache=FALSE,echo=FALSE,fig.height=6>>=
ggplot(data=minutes.data, aes(x=minute,y=PUTs,color=ratio))+ scale_x_datetime()+geom_point()+stat_smooth()
@ %def graph on PUTs per minute
%
\caption{Plotting the contributions per minute in a new and massive
  experiment, as well as the real number as color. Lighter color
  indicates that all PUTs contributed new individuals to the
  experiment, darker color few to none. \label{fig:minute}}
\end{figure}

@@

The dot color in Figure \ref{fig:minute} corresponds to the {\em ratio} or, as indicated above, the actual
number of contributing PUTs, those that have not been discarded due to
repeated sending by clients or (briefly) belonging to the previous
instance of the experiment. It reveals that, in some cases where the
amount of PUTs really increases, they are actually not contributing so
much, while light dots, corresponding to a ratio close to 1, usually
stay in the lowest part of the graphic. 
%
\begin{figure}[ht!b]
\centering
<<entropy, cache=FALSE,echo=FALSE>>=
max <- max(new.experiment.data$IPs)
entropy.lm <- lm( new.experiment.data$milliseconds ~ new.experiment.data$entropy )
exp.palette <- colorRampPalette(brewer.pal(12,"Paired"))
ggplot(data=new.experiment.data)+geom_point(aes(x=entropy,y=milliseconds,size=PUTs,color=IPs))+scale_x_log10()+scale_y_log10()+scale_color_gradient(limits=c(1,max),low='red')
@ %def entropy data
%
\caption{Entropy of the user contribution vs. time to solution in the
  volunteer computing experiment. The fit to a linear model is significant. \label{fig:entropy}}
\end{figure}

>>>>

\begin{document}

<<setup, cache=FALSE,echo=FALSE>>=
library("ggplot2")
library("RColorBrewer")
opts_chunk$set( dev        = c("pdf", "png"),
               fig.width   = 9,
               fig.height  = 6,
               dpi     = 200 )
experiment.data <- data.frame(Experiment=character(),
                              milliseconds=character(),                           
                              IPs=character(), 
                              actualIPs=character(),
                              PUTs=character(),
                              actualPUTs=character(),
                              PUTratio = character(),
                              PUTsIP =character())

@@

%
%
\begin{figure}[htbp]
\centering
<<zipf, cache=FALSE,echo=FALSE, fig.width   = 9, fig.height  = 6>>=
ggplot(ips.data,aes(x=idx,y=PUTs,group=Experiment,color=Experiment))+geom_point()+scale_y_log10()+ scale_colour_discrete(labels=c('base','graph1','graph2','reboot','cachereboot'))+scale_x_log10()
@ %def graph for IPs and puts
%
\caption{Log-log plot of ranked and normalized number of contributions (PUTs) per IP.\label{fig:zipf}}
\end{figure}
%
Every run had slightly different conditions, although in some cases it
was just a server reboot and a new round of announcements. Each run
batch included at least 30 completions, that is, evolution until the
solution was found. The number of runs was controlled by polling an
URL that indicated the number of runs made in the present batch so
far. Every volunteer contributed a certain number of individuals
(represented by the PUT HTTP request in Figure \ref{fig:system}) to the server. The ranked number
of contributions per unique IP is shown in Figure \ref{fig:zipf} in a
log-log scale. Besides following the usual Zipf's law found in our
previous work and thus having a very similar appearance for all 5 sets, it
is interesting to see how some sets have a number of contributions per
IP that are slightly superior to the rest, especially so in {\tt
  graph2}. The last two sets, {\tt reboot} and {\tt cachereboot}, also
show a regime in the middle of the graph which shows that some users
are contributing more than in other similar sets. In general, these
differences in behavior are due more to the specific users that show
up in these experiments than to general laws, but it should also be
noted that those behaviors can be encouraged by design.

@@

\section{Finding the keys to a good performance}
%
%
\begin{figure}[ht!b]
\centering
<<minutedata, cache=FALSE,echo=FALSE, fig.width   = 9, fig.height  = 6, fig.dpi     = 100>>=
ggplot(data=minutes.data, aes(x=minute,y=PUTs,color=ratio))+ scale_x_datetime()+geom_point()+stat_smooth()
@ %def graph on PUTs per minute
%
\caption{Plotting the contributions per minute in a new and massive
  experiment, as well as the real number as color. Lighter color
  indicates that all PUTs contributed new individuals to the
  experiment, darker color few to none. \label{fig:minute}}
\end{figure}

@@

The dot color in Figure \ref{fig:minute} corresponds to the {\em ratio} or, as indicated above, the actual
number of contributing PUTs, those that have not been discarded due to
repeated sending by clients or (briefly) belonging to the previous
instance of the experiment. It reveals that, in some cases where the
amount of PUTs really increases, they are actually not contributing so
much, while light dots, corresponding to a ratio close to 1, usually
stay in the lowest part of the graphic. 
%
\begin{figure}[ht!b]
\centering
<<entropy,cache=FALSE,echo=FALSE,fig.width=9,fig.height=6,dpi=200,dev='png'>>=
max <- max(new.experiment.data$IPs)
entropy.lm <- lm( new.experiment.data$milliseconds ~ new.experiment.data$entropy )
exp.palette <- colorRampPalette(brewer.pal(12,"Paired"))
ggplot(data=new.experiment.data)+geom_point(aes(x=entropy,y=milliseconds,size=PUTs,color=IPs))+scale_x_log10()+scale_y_log10()+scale_color_gradient(limits=c(1,max),low='red')
@ %def entropy data
%
\caption{Entropy of the user contribution vs. time to solution in the
  volunteer computing experiment. The fit to a linear model is significant. \label{fig:entropy}}
\end{figure}
